---
title: "R Notebook"
output: html_notebook
---

```{r}
BreadBasketData <- read.csv("https://raw.githubusercontent.com/reisanar/datasets/master/BreadBasket.csv")
```


```{r}
library(dplyr)
library(tidyverse)
library(arules)
```

Lucy's Work:
I'd like to see if there is a day (M, T, W, R, F, S, S) where more transactions took place and if more transactions were in the morning, afternoon, or evenings.

```{r}
#filter data into three categories: morning, afternoon, evening

MorningData <- BreadBasketData %>% 
                filter(Time <= 10:59:59)   #before 11am

AfternoonData <- BreadBasketData %>% 
                  #filter(Time > 10:59:59 && Time <= 15:59:59)   #11 to 4pm
                  #filter(time, Time("11:00:00":"15:59:59"))

EveningData <- BreadBasketData %>% 
                filter(Time > 16:00:00)   #after 4pm
```


Where to start:
-Summarize the data using any of the tools and concepts discussed so far. For example, you may need to aggregate the data (using filter(), group_by(), and summarize() from the dplyr package). Contact your instructor as soon as possible if you run into trouble formatting the data and performing any data wrangling steps.
-Create summaries, comparisons, and any other type of quantitative analysis using the tools learned in class.
-Briefly describe the methodology you decide to use and explain your choice of descriptive/predictive/prescriptive analytics methods. For example, you may decide to explore clustering methods, association rules, anomaly detection algorithms, topic modeling approaches, sentiment analysis, and/or other data mining and text mining methodologies.
-You should include some data visualizations and enough documentation of the code you develop, noting that a small set of high quality visualizations will receive a much higher grade than a large number of poor quality visualizations.

Chris' Work:

```{r}
#get rid of every column except the transaction and the items
breadbasketmain <- BreadBasketData %>%
  select(Transaction, Item)

#check that it worked
breadbasketmain

#convert new set to transaction
bb_tr <- as(split(breadbasket[ ,"Item"], breadbasket[ ,"Transaction"]), "transactions")
inspect(bb_tr)

#create item frequency plot of the top 6 items
IF_bb = rev(tail(sort(itemFrequency(bb_tr)), 6))
barplot(IF_bb, las=2, cex.names=0.8)
```
